---
title: 'Calculation of fractional VOC losses from silage using 2012 model'
output: pdf_document
author: Sasha D. Hafner
date: 2021 August
---

# Overview
This document presents calculation of fraction VOC losses from silage fed to cattle based on the model described in Hafner et al. (2012). 
All calculations were carried out in R.

# Initial steps
Load model from Hafner et al. (2012).
```{r}
source('../model/FAConvDiffMod_v8.R')
```

Set constants.
Diffusivity in clear air (m2/s) (`d.a`) and Henry's law constant (aq:g, g / m3 in aq phase per g / m3 gas phase).
Henry's law constant parameters from NIST Chemistry Webbook (Sander, 2011). 
```{r}
d.a <- c(AceticAcid = 1.2E-5, Acetaldehyde = 1.3E-5, PropylAcetate = 1.3E-5, Ethanol = 1.2E-5) 

k.h.p <- matrix(c(-12.6,  6300,
                  -17.31, 5920,
                  -17.6,  5700,
                  -15.78, 6248),
              ncol = 2, byrow = TRUE, 
              dimnames = list(c('AceticAcid', 'Acetaldehyde', 'PropylAcetate', 'Ethanol'), 
                              c('int', 'itemp')))
k.h.p
```

# 1. Feedout loss
Assumptions:

* 15 cm removed every 12 hours
* Loss only from outer 15 cm
* Density was 232 kg/m3, based on Muck and Holmes (2000) Table 1 average (WI bunker silos)
* Dry matter 34%, from same source
* Gas-phase diffusion-dispersion coefficient `ksg` is 2E-5 m2/s from Hafner et al. (2012, Fig. 1)
* Temperature was 20 degrees C, close to the 1971-2000 climatological normal average presented in Table 1 in Montes et al. 2010 (packed silage)
* Wind speed is lower than the climatological normal average of about 4 m/s, due to obstructions and resistance close to the ground, giving a surface mass transfer coefficient value of 0.01 m/s

Set silage properties and management.
These follow assumptions listed above.
```{r}
dm <- 0.34
rho.d <- 232
thk <- 0.15
t.outs <- 12 * 3600
```

Transport properties.
```{r}
k.sg <- 2E-5
h.m <- 0.01
```

Weather

```{r}
temp.c <- 20
temp.k <- temp.c + 273.15
```

```{r}
k.h <- exp(k.h.p[, 'int'] + k.h.p[, 'itemp']/temp.k)
k.h
```


## Acetic acid loss
```{r}
cmpd <- 'AceticAcid'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```

Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.

## Ethanol loss
```{r}
cmpd <- 'Ethanol'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```

Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.

## Propyl acetate
```{r}
cmpd <- 'PropylAcetate'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```

Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.

## Acetaldehyde
```{r}
cmpd <- 'Acetaldehyde'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```
Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.

# 2. Feeding loss
Assumptions:

* Average of 20 cm silage in feed lanes for 6 hours
* Density is 75 kg/m3, based roughly on Hafner et al. (2012)
* Dry matter 34%, from Muck and Holmes 2000 Table 1
* Gas-phase diffusion-dispersion coefficient `ksg` is 3.3E-5 m2/s from Hafner et al. (2012, median best-fit value listed on p 139, right side)
* Temperature same as with feedout above
* Surface mass transfer coefficient value of 0.01 m/s from Hafner et al. (2012)

Silage properties and management, based on above assumptions.
```{r}
dm <- 0.34
rho.d <- 75
thk <- 0.15
t.outs <- 6 * 3600
```

Transport properties
```{r}
k.sg <- 3.3E-5
h.m <- 0.01
```

## Acetic acid loss
```{r}
cmpd <- 'AceticAcid'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```

Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.


## Ethanol loss
```{r}
cmpd <- 'Ethanol'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```

Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.

## Propyl acetate
```{r}
cmpd <- 'PropylAcetate'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```

Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.

## Acetaldehyde
```{r}
cmpd <- 'Acetaldehyde'
out <- facd.mod(c.d = 1, d.a = d.a[cmpd], dm = dm, e.d=100, h.m = h.m, k.h = k.h[cmpd],
                l = thk, k.sg = k.sg, temp.c = temp.c, t.outs = t.outs, rho.d = rho.d)
out$ts
```
Predicted loss: `r round(100 * out$ts$f.lost, 1)`%.
